ESBMC-GPU \\ A Context-Bounded Model Checking Tool \\ to Verify CUDA Programs

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
clean up
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

%\vspace{-1mm}

@@

%\textcolor{red}{seria interessante colocar um resumo sobre o que serah visto no resto do artigo?} \textcolor{blue}{Acho que nao temos espaco para isso} \textcolor{green}{Melhor nao colocarmos este resumo...}

@@

%\vspace{-3mm}

@@

%\vspace{-2mm}

\begin{figure}[htb]
  \centering
  \includegraphics[width=4.8in]{../figures/arch.pdf}
  \caption{Overview of ESBMC-GPU's architecture.}
  \label{figure:arch}
\end{figure}

%\vspace{-2mm}

@@

ESBMC was designed to handle multi-threaded software, through the use of an API called Portable Operating System Interface (POSIX -- ISO/IEC 9945)~\cite{posix:2008}. Thus, ESBMC-GPU applies a combination of processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths. Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).\\
%Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.

@@

%\noindent \textbf{State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and reduce state space, based on SHA256 hashes.\\%, due to the low the likelihood regarding collisions. \\
%

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also \textcolor{blue}{reduce the state space}, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.\\% \textcolor{red}{similares? entao, qual eh o impacto de se verificar apenas um? eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.

@@

%In addition to both techniques,

%\vspace{-4mm}

@@

%\textcolor{red}{eu nao sei como a disponibilidade de espaco estah, mas seria bom aqui um fechamento de como BMC checa tudo...} \textcolor{blue}{acredito que nao teremos espaco para isso} \textcolor{green}{Felipe, podes referenciar o nosso journal do CPE informando que la explicamos como o BMC checa tudo com mais detalhes.}

%\vspace{-4mm}

@@

\begin{center}
%\vspace{-2 mm}
\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}

\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
%\vspace{-2 mm}
\end{center}

@@

%\vspace{-0.1cm}

%\textcolor{red}{esta parte ficou solta, pois nao houve explicacao de BMC e, por conseguinte, k (unwind)} \textcolor{blue}{concorco com o comentario, mas como nao temos espaco, estamos considerando que o revisor ja tera conhecimento de tecnicas mais conhecidas, como BMC} \textcolor{green}{sugiro novamente referenciar o journal do CPE.}
%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, respectively. Furthermore, its kernel (lines $8$ to $10$) stores in the variable $c$ the result of the sum of $a$ and $b$, which are passed as input arguments. However, there is a mistake in the assertion (line $26$), as $a + b$ must be equal to $c$, which is correctly detected by ESBMC-GPU detects an assert violation.

%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, and a kernel (lines $12$ to $14$), which assigns thread's index values to an array passed as an input argument. Indeed, its goal is to instantiate array positions according to the thread index. However, there is a mistake in the array index, as the value $1$ is accidentally added to the thread index (line $13$). As shown in the main function, array positions are assigned with value $0$ (line $23$), and after the kernel call (line $26$), it is expected that $a[0] == 0$ and $a[1] == 1$. \textcolor{red}{$<$---- poderiamos deletar esse paragrafo?}

@@

%\vspace{-7mm}

@@

 In order to answer both questions, all benchmarks were verified with $4$ GPU verifiers (ESBMC-GPU $v2.0$, GKLEE $v2012$, GPUVerify $v1811$, and CIVL $v1.7.1$), on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, with $16$ GB of RAM, running Ubuntu 14.04 OS. Importantly, all presented execution times are actually CPU times, {\it i.e.}, only the elapsed time periods spent in the allocated CPUs, which was measured with the {\tt times} system call (POSIX system). An overview of the experimental results is shown in Fig.~\ref{figure:experimental}, where {\it True} represents bug-free benchmarks, {\it False} represents buggy benchmarks, {\it Not supported} represents benchmarks that could not be verified, {\it Correct} represents the percentage of benchmarks correctly verified, and {\it Incorrect} represents the percentage of benchmarks incorrectly verified ({\it i.e.}, a verification tool reports an unexpected result). As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}. More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).\\ %\textcolor{red}{e os outros? quais sao as vantagens? esse paragrafo ficou incompleto} \textcolor{blue}{estamos focando apenas nos aspectos do ESBMC porque nao temos espaco para discutir sobre os demais, isso estara na pagina online} \textcolor{green}{Concordo em colocar o texto explicativo na pagina online e referenciar.}

\begin{figure}[htb]
  \centering
  \includegraphics[width=3.9in]{../figures/experimental-results.pdf}
  \caption{Experimental evaluation of ESBMC-GPU against other verifiers.}
  \label{figure:experimental}
\end{figure}

@@

ESBMC-GPU marks the first application of an SMT-based context-BMC tool that recognizes CUDA directives~\cite{pereira:2016}. \textcolor{blue}{Besides, it also applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.} Finally, it presents improved ability to detect array out-of-bounds and data race violations. Future work aims to support \textcolor{blue}{CUDA (parallel) streams and events~\cite{cuda:2012}} and implement further techniques \textcolor{blue}{(e.g., Abstract Interpretation~\cite{Rocha2017})} to reduce the number of thread interleavings, by taking into account GPU symmetry.

@@

\bibitem{monteiro:2015}
Monteiro, F., Cordeiro, L., and de Lima Filho, E. 
\newblock {\em {Bounded Model Checking of C++ Programs Based on the Qt Framework}}. 
\newblock {GCCE} 2015; 179--447.
%\newblock {DOI=http://dx.doi.org/10.1109/GCCE.2015.7398699}}

\bibitem{garcia:2016}
Garcia, M., Monteiro, F., Cordeiro, L., and de Lima Filho, E. 
\newblock {\em {ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
\newblock {SPIN} 2016; 97--103.
%\newblock {DOI=http://dx.doi.org/10.1007/978-3-319-32582-8\_6}}

>>>>

@@

@@

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=5.5in]{../figures/arch.pdf}
  \caption{Overview of ESBMC-GPU's architecture.}
  \label{figure:arch}
\end{figure}

@@

ESBMC was designed to handle multi-threaded software, through the use of an API called Portable Operating System Interface (POSIX -- ISO/IEC 9945)~\cite{posix:2008}. Thus, ESBMC-GPU applies a combination of processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths. Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).\\

@@

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also \textcolor{blue}{reduce the state space}, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.\\

@@

@@

@@

\begin{center}
\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}

\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
\end{center}

@@

%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, and a kernel (lines $12$ to $14$), which assigns thread's index values to an array passed as an input argument. Indeed, its goal is to instantiate array positions according to the thread index. However, there is a mistake in the array index, as the value $1$ is accidentally added to the thread index (line $13$). As shown in the main function, array positions are assigned with value $0$ (line $23$), and after the kernel call (line $26$), it is expected that $a[0] == 0$ and $a[1] == 1$.

@@

@@

 In order to answer both questions, all benchmarks were verified with $4$ GPU verifiers (ESBMC-GPU $v2.0$, GKLEE $v2012$, GPUVerify $v1811$, and CIVL $v1.7.1$), on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, with $16$ GB of RAM, running Ubuntu 14.04 OS. Importantly, all presented execution times are actually CPU times, {\it i.e.}, only the elapsed time periods spent in the allocated CPUs, which was measured with the {\tt times} system call (POSIX system). An overview of the experimental results is shown in Fig.~\ref{figure:experimental}, where {\it True} represents bug-free benchmarks, {\it False} represents buggy benchmarks, {\it Not supported} represents benchmarks that could not be verified, {\it Correct} represents the percentage of benchmarks correctly verified, and {\it Incorrect} represents the percentage of benchmarks incorrectly verified ({\it i.e.}, a verification tool reports an unexpected result). As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}. More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).\\

\begin{figure}[htb]
  \centering
  \includegraphics[width=4.5in]{../figures/experimental-results.pdf}
  \caption{Experimental evaluation of ESBMC-GPU against other verifiers.}
  \label{figure:experimental}
\end{figure}

@@

ESBMC-GPU marks the first application of an SMT-based context-BMC tool that recognizes CUDA directives~\cite{pereira:2016}. \textcolor{blue}{Besides, it also applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.} Finally, it presents improved ability to detect array out-of-bounds and data race violations. Future work aims to support \textcolor{blue}{CUDA (parallel) streams and events~\cite{cuda:2012}} and implement further techniques \textcolor{blue}{(e.g., Abstract Interpretation~\cite{Rocha2017}) in order to prune the state space exploration}, by taking into account GPU symmetry.

@@

\bibitem{monteiro:2015}
Monteiro, F., Cordeiro, L., and de Lima Filho, E. 
\newblock {\em {Bounded Model Checking of C++ Programs Based on the Qt Framework}}. 
\newblock {GCCE} 2015; 179--447.

\bibitem{garcia:2016}
Garcia, M., Monteiro, F., Cordeiro, L., and de Lima Filho, E. 
\newblock {\em {ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
\newblock {SPIN} 2016; 97--103.
